---
title: "Introduction to nmrec"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to nmrec}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

nmrec is an R package for reading records from a NONMEM control
stream, parsing and modifying records of interest, and writing the
result.

## Reading a control stream

`read_ctl()` reads the NONMEM control stream from the specified path.
Or, if you've already read in the lines as a character vector, you can
use `parse_ctl()`.

```{r}
lines <- nmrec::nmrec_examples[["bayes1"]]
head(lines, 4)

ctl <- nmrec::parse_ctl(lines)
```

The result is an `?nmrec_ctl_records` list of two items, "frontmatter"
(any lines before the first record) and "records".

```{r}
ctl$frontmatter

length(ctl$records)
head(ctl$records, 3)
```

## Record objects

The list of records contains `?nmrec_record` objects.

```{r}
purrr::map_chr(ctl$records, "name")
unique(purrr::map_chr(ctl$records, is))

class(ctl$records[[17]])
```

If nmrec doesn't support parsing _within_ a given record type, the
record type will be `nmrec_record_raw`.  Otherwise, the record type
will be `nmrec_record_{name}`, where `name` is the standardized record
name in lower case.

You can use `select_records()` to extract records of a given type.
This is a simple wrapper around `purrr::keep()` that standardizes the
input name (e.g., "est" to "estimation").

```{r}
ests <- nmrec::select_records(ctl, "est")
head(ests, 2)
```

The `?record` object is an [R6] object.

```{r}
est1 <- ests[[1]]
ls(est1)
```

The primary method of interest is `$parse()`, which parses the record
lines and populates the `$values` field.  This field is a list of
elements (things like spaces and new lines) and `?nmrec_option`
objects.  Once an record is parsed, the `$values` field fully controls
how it will be rendered.

```{r}
est1$values

est1$parse()

length(est1$values)
head(est1$values, 3)
```

## Option objects

You can grab an option of interest from a record with
`get_record_option()`.

```{r}
est1_meth <- nmrec::get_record_option(est1, "meth")
ls(est1_meth)
```

`?nmrec_option` objects, like `?nmrec_record` objects, are [R6]
objects.  The fields of interest depend on the option type.  Here are
the important fields for the selected method option of the first
estimation record:

```{r}
est1_meth$name
est1_meth$name_raw
est1_meth$value
est1_meth$sep
```

## Modifying options and records

A key feature of option and records objects is that, as [R6] objects,
they have **reference semantics** (see `vignette("Introduction",
"R6")`).  For example, you could mutate `NSIG=2` to `NSIG=3` in the
`est1` record like so:

```{r}
est1_nsig <- nmrec::get_record_option(est1, "sig")
est1_nsig$name
est1_nsig$value
est1_nsig$value <- 3

est1_nsig
est1
```

If you print `ctl` (the original `nmrec_ctl_records` object), you'll
notice that all of the changes above are reflected there.  You can
write the changes to a file with `write_ctl()`.

```{r, eval = FALSE}
write_ctl(ctl, "modified.ctl")
```

[R6]: https://r6.r-lib.org/
